---
title: "Analysis of Readiness Examination Scores (System-Wide)"
date: "10/30/2018"
output:
  pdf_document: default
  word_document: default
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 8, fig.align = "center", cache = TRUE, ggplot2::theme_set(ggplot2::theme_bw()))
```

```{r}
source("re_data_structuring.R")
```

## Executive Summary

* The MCQ RE scores for Grade 8 continue to improve in general with the exception of Integrated Science, where the scores are lower compared to the 2017 scores.
* There are remarkable improvements in the Grade 8 non-MCQ RE scores for Filipino, IS, and Math. However, there is a continous downward trend in the scores for Social Science.
* The Grade 10 MCQ scores for English follow an increasing trend. There are remarkable increases in the scores for Filipino and Social Science. On the other hand, there is an appreciable decrease in the Physics scores.
* In terms of the non-MCQ tests, only English has had a remarkable increase in the distribution of scores. 
* In general, the scores for the MCQ tests are normally distributed. This means that, in the MCQ tests, about 68% of the students scored 1 standard deviation about the mean score. Visible skewness in the distribution of scores can be found in some of the Non-MCQ scores. Notably:
  - in Grade 8 Filipino Non-MCQ test, more students got scores that are higher than the mean score;
  - in Grade 8 Social Science and English Non-MCQ tests, more students got scores that are lower than the mean score;
  - an appreciable number of students scored lower than the mean scores in the Grade 10 Chemistry and Math Non-MCQ tests.
* In the following analyses, 50% is used to benchmark the scores of the students. There are two ways of viewing this benchmark: the subjects are easy or the students are not "ready" in more than 50% of the skills covered in the tests. These two views may not be mutually exclusive of each other. Also, the 50% benchmark is arbitrary and can be set by the subject-matter experts. However, there is the danger of setting the benchmark too low in order to make it appear that the students performed well when in fact they fail to show readiness in most skills required by the items that they fail to answer correctly[^1]. 
  - In terms of the MCQ test scores for Grade 8, more than half of the students who took the tests got more than 50% of the items correctly. However, more than 2 out of 5 students scored below 50% in Integrated Science. This is followed by Math, in which 29% of the students scored below 50%. On the other hand, 97% of the students scored above 50% in English.
  - In terms of the non-MCQ tests for Grade 8, only 30% of the test-takers scored above 50% in Social Science, which is the lowest among all of the subjects. On the other hand, 92% of the test-takers got at least 50% of the items correctly in the Filipino non-MCQ test.
  - More than 50% of the Grade 10 test-takers scored 50% or above in all Grade 10 MCQ subjects except for Math (47%) and Physics (34%). However, Statistics (52%) and Chemistry (60%) also have noticeably low percentages of test-takers that scored above 50%.
  - In all the non-MCQ tests in STEM, the test-takers got scores below 50%. More worrisome is the fact that in the computational subjects, less than 20% got scores above 50% (Math---9%; Stat---10%; Physics---19%). However, this is not surprising as this has been the trend for these subjects in the past years.
  - It is also not surprising that the students performed worse in the STEM subjects than in the Humanities subjects, overall. But the gap in performance is alarming, nonetheless.
  - The students also performed worse in Non-MCQ tests in general. This difference is remarkably noticeable for STEM subjects. Also, scores in Humanities subjects varied more in the non-MCQ tests than in the MCQ tests.

[^1]: Morever, the subject-matter experts can also identify critical skills that are required for the higher years in PSHS. But this rather complicates how we assess the readiness of our students for higher learning. In the end, we should strive for a balance between the ability to assess the nuances of our students' readiness for more complicated tasks and the expediency of determining such readiness for intervention purposes.
  

## Statistics in Figures


```{r}
re2015_18 <- setDT(re2015_18)[,stype := case_when(subject %in% c("english", "filipino", "ss")~"Humanities",
                              TRUE~"STEM")]
```

```{r, include=FALSE}
re2015_18 <- re2015_18[assessment == "re"][,`:=`(total = sum(score, na.rm = T), total_max = sum(maxscore, na.rm = T)), by = .(year, grade, subject, sname)]
```

```{r}
re2015_18[,pct := round(score / maxscore * 100, 2)]
re2015_18[,pct_tot := round(total / total_max * 100, 2)]
```

### Distribution of Grade 8 MCQ scores from 2015-2018

```{r, fig.height = 3}
re2015_18[test_type == "mcq" & total > 0 & score > 0 & grade == "8", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(year, pct, group = year)) + geom_boxplot() + facet_wrap(~subject, nrow = 1) +
  ylab("Percent") +
  xlab("Subject")
```

### Distribution of Grade 10 MCQ scores from 2015-2018

```{r, fig.height=6}
re2015_18[test_type == "mcq" & total > 0 & score > 0 & grade == "10", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(year, pct, group = year)) + geom_boxplot() + facet_wrap(~subject, nrow = 2) +
  ylab("Percent") +
  xlab("Subject")
```

### Distribution of Grade 8 Non-MCQ scores from 2015-2018

```{r, fig.height = 3}
re2015_18[test_type == "non-mcq" & total > 0 & score > 0 & grade == "8", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(year, pct, group = year)) + geom_boxplot() + facet_wrap(~subject, nrow = 1) +
  ylab("Percent") +
  xlab("Subject")
```

### Distribution of Grade 10 Non-MCQ scores from 2015-2018


```{r, fig.height=6}
re2015_18[test_type == "non-mcq" & total > 0 & score > 0 & grade == "10", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(year, pct, group = year)) + geom_boxplot() + facet_wrap(~subject, nrow = 2) +
  ylab("Percent") +
  xlab("Subject")
```


### Distribution of Grade 8 MCQ scores for 2018 RE

```{r, fig.height=3}
summaryre2018g8 <- re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "8", .(mean = round(mean(pct)), SD = round(sd(pct)), median = round(median(pct))), by = .(year, grade, subject)][,subject := toupper(subject)]
re2015_18[test_type == "mcq" & total > 0 & score > 0 & grade == "8" & year == "2018", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(pct, group = year)) + geom_histogram(aes(y = 100*..density..), bins = 18, alpha = 0.7, color = "white") + facet_wrap(~subject, nrow = 1) + geom_density(aes(y = 100 *..density..)) +
  geom_vline(data = summaryre2018g8, aes(xintercept = mean), color = "red") +
  geom_vline(data = summaryre2018g8, aes(xintercept = median), color = "blue") +
  geom_text(data = summaryre2018g8, aes(x = 20, y = 5.5, label = paste("Mean = ", mean)), nudge_x = 10, color = "red") +
  geom_text(data = summaryre2018g8, aes(x = 20, y = 5, label = paste("Median = ", median)), nudge_x = 10, color = "blue") +
  xlab("Score (%)") +
  ylab("Percentage of Test Takers")
```

### Distribution of Grade 10 MCQ scores for 2018 RE

```{r, fig.height=6}
summaryre2018g10 <- re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10", .(mean = round(mean(pct)), SD = round(sd(pct)), median = round(median(pct))), by = .(year, grade, subject)][,subject := toupper(subject)]
re2015_18[test_type == "mcq" & total > 0 & score > 0 & grade == "10" & year == "2018", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(pct, group = year)) + geom_histogram(aes(y = 100*..density..), bins = 18, alpha = 0.7, color = "white") + facet_wrap(~subject, nrow = 2) + geom_density(aes(y = 100 *..density..)) +
  geom_vline(data = summaryre2018g10, aes(xintercept = mean), color = "red") +
  geom_vline(data = summaryre2018g10, aes(xintercept = median), color = "blue") +
  geom_text(data = summaryre2018g10, aes(x = 20, y = 5.5, label = paste("Mean = ", mean)), nudge_x = 10, color = "red") +
  geom_text(data = summaryre2018g10, aes(x = 20, y = 5, label = paste("Median = ", median)), nudge_x = 10, color = "blue") +
  xlab("Score (%)") +
  ylab("Percentage of Test Takers")
```


### Distribution of Grade 8 Non-MCQ scores for 2018 RE

```{r, fig.height=3}
summaryre2018g8nmcq <- re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "8", .(mean = round(mean(pct)), SD = round(sd(pct)), median = round(median(pct))), by = .(year, grade, subject)][,subject := toupper(subject)]
re2015_18[test_type == "non-mcq" & total > 0 & score > 0 & grade == "8" & year == "2018", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(pct, group = year)) + geom_histogram(aes(y = 100*..density..), bins = 10, alpha = 0.7, color = "white") + facet_wrap(~subject, nrow = 1) + geom_density(aes(y = 100 *..density..)) +
  geom_vline(data = summaryre2018g8nmcq, aes(xintercept = mean), color = "red") +
  geom_vline(data = summaryre2018g8nmcq, aes(xintercept = median), color = "blue") +
  geom_text(data = summaryre2018g8nmcq, aes(x = 20, y = 5.5, label = paste("Mean = ", mean)), nudge_x = 10, color = "red") +
  geom_text(data = summaryre2018g8nmcq, aes(x = 20, y = 5, label = paste("Median = ", median)), nudge_x = 10, color = "blue") +
  xlab("Score (%)") +
  ylab("Percentage of Test Takers")
```

### Distribution of Grade 10 Non-MCQ scores for 2018 RE

```{r, fig.height=6}
summaryre2018g10nmcq <- re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10", .(mean = round(mean(pct)), SD = round(sd(pct)), median = round(median(pct))), by = .(year, grade, subject)][,subject := toupper(subject)]
re2015_18[test_type == "non-mcq" & total > 0 & score > 0 & grade == "10" & year == "2018", ] %>% mutate(year = lubridate::ymd(year, truncated = 2L)) %>%
  mutate(subject = toupper(subject)) %>%
  ggplot(aes(pct, group = year)) + geom_histogram(aes(y = 100*..density..), bins = 7, alpha = 0.7, color = "white") + facet_wrap(~subject, nrow = 2) + geom_density(aes(y = 100 *..density..)) +
  geom_vline(data = summaryre2018g10nmcq, aes(xintercept = mean), color = "red") +
  geom_vline(data = summaryre2018g10nmcq, aes(xintercept = median), color = "blue") +
  geom_text(data = summaryre2018g10nmcq, aes(x = 20, y = 5.5, label = paste("Mean = ", mean)), nudge_x = 10, color = "red") +
  geom_text(data = summaryre2018g10nmcq, aes(x = 20, y = 5, label = paste("Median = ", median)), nudge_x = 10, color = "blue") +
  xlab("Score (%)") +
  ylab("Percentage of Test Takers")
```

### Distribution of Scores: Humanities vs STEM
```{r, fig.height=3}
re2015_18[year == "2018"] %>%
  ggplot(aes(stype, pct)) +
  geom_boxplot() +
  facet_wrap(~test_type)
```

## Statistics in Numbers

```{r}
library(pander)
library(moments)
```

### Grade 8 MCQ Test

```{r}
below50 <- function(x) {x >= 50}
stats <- function(x) {
  list(
    N = length(x),
    mean = round(mean(x),2),
    SD = round(sd(x),2),
    min = min(x),
    Q1 = quantile(x, .25),
    median = round(median(x)),
    Q3 = quantile(x, .75),
    max = max(x),
    `% >= 50` = round(sum(as.integer(below50(x))) / length(x) * 100)
    )
}

re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "8", stats(pct), by = .(year, grade, subject)][, subject := toupper(subject)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "8", stats(pct), by = .(year, grade, subject)][, subject := toupper(subject)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```


### Grade 10 MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10", stats(pct), by = .(year, grade, subject)][, subject := toupper(subject)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```


### Grade 10 Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10", stats(pct), by = .(year, grade, subject)][, subject := toupper(subject)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

## Relative Performances of Campus

### Grade 8 English MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "8" & subject == "english", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 English Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "8" & subject == "english", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Filipino MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "8" & subject == "filipino", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Filipino Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "8" & subject == "filipino", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Integrated Science MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "8" & subject == "is", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Integrated Science Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "8" & subject == "is", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Mathematics MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "8" & subject == "math", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Mathematics Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "8" & subject == "math", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```


### Grade 8 Social Science MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "8" & subject == "ss", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 8 Social Science Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "8" & subject == "math", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Biology MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "bio", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```


### Grade 10 Biology Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "bio", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Chemistry MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "chem", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Chemistry Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "chem", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 English MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "english", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 English Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "english", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Filipino MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "filipino", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Filipino MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "filipino", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Mathematics MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "math", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Mathematics Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "math", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Physics MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "phys", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Physics Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "phys", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Social Science MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "ss", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Social Science Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "ss", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Statistics MCQ Test

```{r}
re2015_18[test_type == "mcq" & score > 0 & year == "2018" & grade == "10" & subject == "stat", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```

### Grade 10 Statistics Non-MCQ Test

```{r}
re2015_18[test_type == "non-mcq" & score > 0 & year == "2018" & grade == "10" & subject == "stat", stats(pct), by = .(year, grade, campus)] %>% select(-1,-2) %>% arrange(desc(`% >= 50`)) %>% pander()
```