---
title: "Association of RE scores and the campuses's involvement in the RE TWG: Grade 10"
author: "Joseph S. Tabadero, Jr."
date: "November 1, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning =F, message = F, fig.height = 7, fight.width = 12, ggplot2::theme_set(ggplot2::theme_bw()), out.width = "55%", fig.align = "center")
```

## Summary of Results

* In general, there is no association between RE MCQ and non-MCQ scores and a campus's involvement in the RE TWG.
* The Grade 10 MCQ scores have a positive association with a campus's involvement in the TWG in 2016, except for Statistics. But this association can be attributed to the concentration of TWG members from perenially top performing campuses such as Main Campus, Bicol Region Campus, Central Luzon Campus, and Southern Mindanao Campus. The same is also true for the positive association found in the Grade 8 RE MCQ scores for 2015 and 2016.


```{r}
source("re_data_structuring.R")
```

```{r}
twg <- fread("./Data/processed/twg.csv", na = "")
colnames(twg) <- tolower(colnames(twg))
twg <- twg %>%
  mutate(subject = tolower(case_when(subject %in% c("integrated", "Integrated", "Integrated Science")~"is",
                             subject == "Mathematics"~"math",
                             subject == "Biology"~"bio",
                             subject == "Chemistry"~"chem",
                             subject == "Physics"~"phys",
                             subject == "Social Science"~"ss",
                             subject == "Statistics"~"stat",
                             TRUE~subject))) %>%
  mutate(grade = as.integer(str_extract_all(grade, "[0-9]+"))) %>%
  mutate(campus = str_remove_all(campus, "PSHS-"))
```

```{r}
retwg <- left_join(re2015_18, twg, by = c("campus", "year", "subject")) %>%
  rename(grade = grade.x) %>% filter(assessment == "re")
retwg <- retwg %>%
  mutate(withtwg = case_when(
    (year %in% c("2015", "2016")) & !is.na(name)~"with twg",
    (year %in% c("2015", "2016")) & is.na(name)~"no twg",
    (year %in% c("2017", "2018")) & is.na(name)~"no twg",
    TRUE ~ twg,
    )) %>%
  mutate(withtwg2 = case_when(
    !is.na(name)~"with twg",
    TRUE~"no twg"
  )
           ) %>%
  mutate(pct = round(score / maxscore * 100, 2))
```

## Distribution of TWG members by grade level, campus, and year

```{r}
twg %>% tabyl(grade, campus, year) %>% enframe() %>% unnest() %>% slice(3:n()) %>%
  mutate(grade = replace_na(grade, 8)) %>%
  rename(Year = name, Grade = grade) %>%
  select(-Year) %>%
  kable("latex", booktabs = TRUE) %>% group_rows(group_label = "2015", start_row = 1, end_row = 1) %>%
  group_rows("2016", start_row = 2, end_row = 3) %>% group_rows("2017", start_row = 4, end_row = 5) %>%
  group_rows("2018", start_row = 6, end_row = 7) %>% kable_styling(latex_options = "scale_down")
```


## Grade 8 System-wide MCQ Scoress

### 2018

```{r, fig.height=4}
retwg %>% filter(test_type == "mcq", grade == "8", year == "2018") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5)
```

### 2017

```{r, fig.height=4}
retwg %>% filter(test_type == "mcq", grade == "8", year == "2017") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5)
```

### 2016

```{r, fig.height=4}
retwg %>% filter(test_type == "mcq", grade == "8", year == "2016") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5)
```

### 2015

```{r, fig.height=4}
retwg %>% filter(test_type == "mcq", grade == "8", year == "2015") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5)
```









## Grade 8 Social Science MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "ss", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "ss", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "ss", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```
### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "ss", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```
## Grade 8 English MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```
## Grade 8 Filipino MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "filipino", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "filipino", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "filipino", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 8 Mathematics MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "math", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip()  + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "math", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```



### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "math", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```




### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "math", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```



## Grade 8 System-wide Non-MCQ Scoress

### 2018

```{r, fig.height=4}
retwg %>% filter(test_type == "non-mcq", grade == "8", year == "2018") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5) + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=4}
retwg %>% filter(test_type == "non-mcq", grade == "8", year == "2017") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5)
```

### 2016

```{r, fig.height=4}
retwg %>% filter(test_type == "non-mcq", grade == "8", year == "2016") %>%
  ggplot(aes(withtwg, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5)
```


### 2016

```{r, fig.height=4}
retwg %>% filter(test_type == "non-mcq", grade == "8", year == "2015") %>%
  ggplot(aes(withtwg, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 5)
```





## Grade 8 Social Science non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "ss", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "ss", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "ss", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "ss", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 8 English non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 8 Filipino non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "filipino", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "filipino", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() +scale_fill_brewer(palette="Greys")
```

### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "filipino", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() +scale_fill_brewer(palette="Greys")
```

## Grade 8 Mathematics non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "math", grade == "8", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "math", grade == "8", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```



### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "math", grade == "8", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2015

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "math", grade == "8", year == "2015") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 System-wide MCQ Scoress

### 2018

```{r}
retwg %>% filter(test_type == "mcq", grade == "10", year == "2018") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 4)
```

### 2017

```{r}
retwg %>% filter(test_type == "mcq", grade == "10", year == "2017") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 4)
```

### 2016

```{r}
retwg %>% filter(test_type == "mcq", grade == "10", year == "2016") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 4)
```

## Grade 10 Physics MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "phys", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "phys", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "phys", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```



## Grade 10 Statistics MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "stat", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "stat", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "phys", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Social Science MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "ss", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "ss", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "ss", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 English MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Filipino MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "english", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "filipino", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "filipino", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


## Grade 10 Chemistry MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "chem", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "chem", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "chem", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Biology MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "bio", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "bio", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "bio", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Mathematics MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "math", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "math", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```



### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "mcq", subject == "math", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```







## Grade 10 System-wide Non-MCQ Scoress

### 2018

```{r}
retwg %>% filter(test_type == "non-mcq", grade == "10", year == "2018") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 4)
```

### 2017

```{r}
retwg %>% filter(test_type == "non-mcq", grade == "10", year == "2017") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 4)
```

### 2016

```{r}
retwg %>% filter(test_type == "non-mcq", grade == "10", year == "2016") %>%
  ggplot(aes(withtwg2, pct)) + geom_boxplot() +
  facet_wrap(~grade + subject, ncol = 4)
```

## Grade 10 Physics non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "phys", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "phys", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "phys", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```



## Grade 10 Statistics non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "stat", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "stat", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "phys", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Social Science non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "ss", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "ss", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "ss", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 English non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Filipino non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "english", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "filipino", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "filipino", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


## Grade 10 Chemistry non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "chem", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "chem", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "chem", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Biology non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "bio", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "bio", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```


### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "bio", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

## Grade 10 Mathematics non-MCQ Scores

### 2018

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "math", grade == "10", year == "2018") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

### 2017

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "math", grade == "10", year == "2017") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```



### 2016

```{r, fig.height=3, fig.width=5}
retwg %>% filter(test_type == "non-mcq", subject == "math", grade == "10", year == "2016") %>%
  group_by(campus) %>%
  mutate(mean = mean(pct)) %>%
  ungroup() %>%
  mutate(campus = fct_reorder(campus, mean, .desc = TRUE)) %>%
  ggplot(aes(campus, pct, fill = withtwg)) + geom_boxplot() +
  coord_flip() + scale_fill_brewer(palette="Greys")
```

