---
title: "On Expanding the Enrolment in the Campuses under the PSHS System"
author: Joseph S. Tabadero, Jr.\thanks{Special Science Teacher IV, Philippine Science
  High School-CAR Campus}
date: "9/6/2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    df_print: paged
classoptions: a4paper, 11pt
toc: yes
header-includes:
  - \input{preamble.tex}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning =F, cache = T)
```


```{r}
source("nce_data_structuring.R")
```

## Executive Summary

* Expand the enrolment in the following campuses:

  - Main Campus
  - CLC
  - SMC
  - WVC

* Intensify efforts to increase the quality and number of applicants for MRC, SRC, and ZRC.

## Philippine Science High School Everywhere

With the establishment of the MIMAROPA campus, the Philippine Science High School is now one region away from establishing a campus in every region in the country---as mandated by Republic Act 8496 (otherwise known as "Philippine Science High School (PSHS) System Act of 1997") as amended by Republich Act 9036. This mandate empowers the PSHS System's Board of Trustees to "develop policies for the expansion of enrollment in campuses under the PSHS System" (Section 8 (j)). Although not specified in the Implementing Rules and Regulations of RA 9036, the IRR of RA 8496 identified four factors upon which the expansion of student enrolment shall be based (4.10):

  * Availability of budgetary support
  * Availability of highly qualified faculty members
  * Maintenance of desirable student-teacher ratio
  * Availability of laboratory and other school facilities

One can conflate (1) the first and fourth factors and (2) the second and third factors. It can be argued that budgetary support can be made available if there is a valid need for such support. The expansion of scholarship slots in the CALABARZON Campus, for one, was based on the number of principal qualifiers. On top of this consideration, we shall look at the the number of principal qualifiers from each region who made it to the top 1620 NCE passers in the National Competitive Examinations (NCE). This choice of measure for determining whether or not enrolment in a campus should be expanded is based on the assumption that if the quality of principal qualifiers are the same throughout the country, then the principal qualifiers from each campus should belong to the top 1620 NCE passers in the country.

## Counting Qualified Applicants

While the PSHS System is mandated to provide equal opportunity for each region to have an access to a PSHS campus, RA 8496 also requires standards for it to follow by "offering scholarships to deserving students" (Title I, Sec. 4). This is done through a rigorous selection process starting with the National Competitive Examination (NCE). 

The NCE is composed of 4 tests---verbal, mathematics, science, and abstract reasoning. The "Criteria and Procedures for Selection of Principal and Alternate Qualifiers" lay down the conditions that have to be met by NCE applicants to principal or alternate qualifiers in the PSHS Campuses. 

However, for this paper, we shall only be concerned with an NCE passer and alternate passer. For the purposes of this paper, an *NCE passer* is defined as any NCE applicant whose score in each subject is above the subject mean score for the applicable NCE year.  Also, for the purposes of this paper, we shall take the following definition of an *alternate passer*:

> To be an alternate in the Main Campus, an examinee should satisfy the following eligibility:

> i. Total score is above the national mean; and

> ii. Sub-test Scores in Quantitative and Scientific Abilities must all be greater than or equal to the corresponding national means. On the other hand Sub-test Scores in Verbal Aptitude and Abstract Reasoning must be greater than or equal to 90% of the national mean. For example, if the national mean score in verbal aptitude is 25, an examinee is eligible as alternate if his/her score is at least 22.50.

> To be an alternate in the Regional Campus, an examinee should satisfy the following eligibility:

> i.	Total score is above the national mean; and 

> ii.	Sub-test Score per subject must be greater than or equal to 90% of the national mean. For example, if the national mean score in verbal aptitude is 25, an examinee is eligible as alternate if his/her score is at least 22.50. 


The reason why we are looking at the NCE passers and the alternate passers as a whole is that principal and alternative qualifiers are determined by factors such as campus of preference (first and second) and the number of already existing scholarship slots in the campuses. A less obvious---but more standard---indicator is *the number of passers each campus (as first choice) has in the top 1620 passing examinees nationwide*. Here, we note that a principal qualifier in a campus might not be one of the top 1620 passers. We arrive at the number 1620 by assuming 240 scholarship slots for Main Campus, 120 scholarship slots for CALABARZON Region Campus, and 90 scholarship slots for each of the other 14 existing regional campuses. Our assumption is that if the quality of applicants are the same throughout the regions of the country, then the number of NCE passers in the top 1620 from each region should be the same as the corresponding scholarship slots, which should be primarily dictated by the population of that region. This indicator can then show us: (i) how many NCE passers who belonged to the top 1620 did not qualify as a principal qualifier; and, alternatively, (ii) how many principal qualifiers did not belong to the top 1620 NCE passers. Counting the number of the top 1620 NCE passers in each campus gives us an idea of whether or not we are giving scholarship slots to the most deserving applicants, while maintaining equal opportunity for applicants from the different regions to avail of a scholarship.

The top 3180 NCE passers, on the other hand, can give us an idea on how deep in the list of alternate qualifiers the PSHS can go and still ensure that the quality of scholars is not compromised. Again, we can count the number of passers for each campus of choice who belong to the top 3180 NCE passers.

## Why do we have identify the number of top 1620 passers?

Our attempt to count the number of passers who belong to the top 1620 NCE Examinees will make more sense if we count the number of passers and possible alternates per year. In Table 1, we can see that we have many passers than available scholarships throughout the years. However, the campuses differ in terms of preference by the NCE applicants. For instance, Main Campus has been perennially chosen by many applicants as their campus of first choice. In the other extreme, we see very few passers choosing MIMAROPA Regional Campus since 2015. Since only a limited number of scholarship slots are available, a system of choosing the principal qualifiers and alternate qualifiers is deviced. This system ensures the equitable distribution of scholarship slots per region. For instance, in the event that a passer does not qualify as a principal qualifier in his/her campus of first choice (usually Main Campus), he/she can opt to grab the scholarship slot being offered in his/her campus of second choice (usually in his/her home region). 

The distribution of passers and possible alternate qualifiers by region of origin are shown in Table 2. From this table, we can see that throughout the years, MIMAROPA has the least number of passers, followed by Zamboanga Peninsula.

While the present system ensures the equitable distribution of scholarship slots throughout the regions of the country on the one hand, it begs the question of whether the scholarship slots are being given to the best (in terms of NCE total score and sub-test scores) among the passers in the entire country.

Table 3 shows the number of passers (N) belonging to the top 1620 and the number of passers not belonging to the top 1620 number of passers (Diff) per campus by year since 2008. Considering the Diff values, we have the following realizations:
  
  * Diff values that are close to zero means that almost if not all of the campus's principal qualifiers belong to the top 1620 NCE passers. This means that all of the students are expected to perform very well under the PSHS curriculum.
  * Diff values that are positive and well above zero imply that the campus has more than its share of principal qualifiers in the top 1620 NCE passers. These campuses may offer more scholarship slots with guarantee that the additional slots will be given to deserving students in terms of aptitude. Increasing the number of scholarship slots in such campuses is recommended.
  * Diff values that are negative and well below zero imply that the campus has less than its share of principal qualifiers in the top 1620 NCE passers. This means that there are other scholars, perhaps from other regions, that are more deserving, in terms of aptitude, to be given the scholarhsip slots. However, it may be the case that the quality of instruction of these campuses can appreciably uplift the performance of every student. This factor should be investigated further in order to determine when to increase the scholarship slots in such campuses.

```{r}
kable(APF_by_campus_year, format = "latex", booktabs = TRUE, 
      caption = "Number of Passers and Possible Alternate Qualifiers by Regional Campus",
      col.names = c("Campus",rep(c("P", "A"), times = 10))) %>% kable_styling(latex_options = "scale_down") %>%
  add_header_above(c(" "=1,"2008"=2, "2009"=2, "2010"=2, "2011"=2, "2012"=2,"2013"=2,"2014"=2,"2015"=2,"2016"=2,"2017"=2)) %>%
  add_footnote(c("For CBZRC, CRC, MRC, SRC, and ZRC, the zeros indicate that the campuses were not yet operational during those years.", "The NA under campus is for applicants who did not signify a campus of choice."))
```




```{r}
kable(regional %>% select(-1), format = "latex", booktabs = TRUE, 
      format.args = list(big.mark = ","),
      caption = "Number of Passers (P), Possible Alternate Qualifiers (A) and Number of Promoted Grade 6 Students (Grade 6) by Region from 2008 to 2017",
      col.names = c("Region",rep(c("P", "A", "Grade 6"), times = 10))) %>% kable_styling(latex_options = "scale_down") %>%
  add_header_above(c(" "=1,"2008"=3, "2009"=3, "2010"=3, "2011"=3, "2012"=3,"2013"=3,"2014"=3,"2015"=3,"2016"=3,"2017"=3)) %>%
  landscape() %>%
  add_footnote(c("Data on promoted Grade 6 students taken from the Department of Education."))
```

```{r}
ncetop1620 <- ncetd %>% 
  filter(passflag2 == "P") %>%
  group_by(year) %>%
  top_n(1620, score)
```

```{r}
# ncetd %>% tabyl(year, pass_flag, passflag2)
```
```{r}
ncetop1620diff <- ncetop1620 %>% tabyl(choice1, year) %>%
  gather(year, N, 2:11) %>% 
  mutate(Diff = case_when(choice1 == "MC"~N-240,
                          choice1 == "CBZRC"~N-120,
                          TRUE ~ N-90)) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))
ncetop1620diffwide <- dcast(setDT(ncetop1620diff), choice1~year, value.var = c("N", "Diff"), fun.aggregate = sum)
fwrite(ncetop1620diffwide, "ncetop1620diffwide.csv")
```

```{r}
ncetop1620reg <- ncetop1620 %>% tabyl(choice1, year) %>%
  gather(year, N, 2:11) %>% 
  mutate(Diff = case_when(choice1 == "MC"~N-240,
                          choice1 == "CBZRC"~N-120,
                          TRUE ~ N-90)) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))
ncetop1620diffwide <- dcast(setDT(ncetop1620diff), choice1~year, value.var = c("N", "Diff"), fun.aggregate = sum)
fwrite(ncetop1620diffwide, "ncetop1620diffwide.csv")
```


```{r}
#ncetop1620diff %>%
#  unite(Diff, N, Diff, sep = ": ") %>%
#  spread(year, Diff) %>%
#  rename(Campus = choice1, "2008"= "2011"="2011-01-01", "2012"="2012-01-01", "2013"="2013-01-01", "2014"="2014-01-01", "2015"="2015-01-01", "2016"="2016-01-01", "2017"="2017-01-01") %>%
#  filter(!is.na(Campus)) %>%
#  kable("latex", booktabs = TRUE, caption = "Number of students belonging to the top 1590 examinees by signified campus of first choice, $N_1$, and the number of students who did not make it as a principal qualifier in their choice campus (positive) or the number of students who did not make it to the top 1590 nationally but made it as principal qualifiers (negative) by region, $N_2$ ($N_1: N_2$).")
```

```{r, fig.width=12, fig.height=7}
ncetop1620diff %>% filter(!is.na(choice1)) %>% ggplot(aes(year, Diff)) + geom_line() + geom_hline(aes(yintercept = 0), linetype = "dashed") + facet_wrap(~choice1, scales = "free_y") + 
  xlab("Year") +
  theme_bw()
```



```{r}
ncetop3240 <- ncetd %>% 
  filter(passflag2 %in% c("P", "A")) %>%
  group_by(year) %>%
  top_n(3180, score)
```

```{r}
# ncetd %>% tabyl(year, pass_flag, passflag2)
```
```{r}
ncetop3240diff <- ncetop3240 %>% tabyl(choice1, year) %>%
  gather(year, N, 2:11) %>% 
  mutate(Diff = case_when(choice1 == "MC"~N-240,
                          choice1 == "CBZRC"~N-120,
                          TRUE ~ N-90)) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))
ncetop3240diffwide <- dcast(setDT(ncetop3240diff), choice1~year, value.var = c("N", "Diff"), fun.aggregate = sum)
fwrite(ncetop3240diffwide, "ncetop3240diffwide.csv")
```

```{r}
ndiff <- fread("ndiff.csv", na.strings = "")

ndiff %>%
  mutate(Campus = case_when(is.na(Campus)~"No Choice",
                            TRUE~Campus)) %>%
  kable("latex", 
        booktabs = T,
        col.names = c("Campus", rep(c("N", "Diff"), times = 10)),
        caption = "Number of students belonging to the top 1620 examinees by signified campus of first choice, $N$, and the number of students who did not make it as a principal qualifier in their choice campus (positive) or the number of students who did not make it to the top 1620 nationally but made it as principal qualifiers (negative) by region, Diff.") %>%
  kable_styling(latex_options = "scale_down") %>%
  add_header_above(c(" "=1, "2008"=2,"2009"=2,"2010"=2,"2011"=2,"2012"=2,"2013"=2,"2014"=2,"2015"=2,"2016"=2,"2017"=2))
```

```{r}
ndiff3240 <- fread("ndiff3240.csv", na.strings = "")

ndiff3240 %>%
  mutate(Campus = case_when(is.na(Campus)~"No Choice",
                            TRUE~Campus)) %>%
  kable("latex", 
        booktabs = T,
        col.names = c("Campus", rep(c("N", "Diff"), times = 10)),
        caption = "Number of passers and alternate passers belonging to the top 3240 examinees by signified campus of first choice, $N$, and the number of students who did not make it as a principal qualifier in their choice campus (positive) or the number of students who did not make it to the top 3240 nationally but made it as principal qualifiers (negative) by region, Diff.") %>%
  kable_styling(latex_options = "scale_down") %>%
  add_header_above(c(" "=1, "2008"=2,"2009"=2,"2010"=2,"2011"=2,"2012"=2,"2013"=2,"2014"=2,"2015"=2,"2016"=2,"2017"=2))
```


```{r, fig.width=12, fig.height = 7}
ncetop3240diff %>% filter(!is.na(choice1)) %>% ggplot(aes(year, Diff)) + geom_line() + geom_hline(aes(yintercept = 0), linetype = "dashed") + facet_wrap(~choice1, scales = "free_y") +
  xlab("Year") +
  theme_bw()
```

```{r}
#ncetop3180 %>%
#  group_by(year, choice1, id) %>%
#  summarise(total = sum(score)) %>%
#  ggplot(aes(total)) + geom_histogram(bins = 30) + facet_wrap(~choice1)
```


```{r}
# uncomment in final report
# ncetop3180 %>%
#   group_by(choice1, year, id) %>%
#   summarise(total = sum(score)) %>%
#   mutate(pct = total / 220 * 100) %>%
#   summarise_at(vars(pct), funs(min,q1 = quantile(.,0.25), median, mean, q3 = quantile(.,.75),max, sd)) %>%
#   kable("markdown", latex_options = c("striped", "repeat_header"), digits = 2, caption = "Distribution of Total Score Percentages by Campus and by NCE Year")
```

We have the following observations from table 3:

1. Since 2011, CLC has an average of 120 NCE passers who belong to the top 1620 NCE passers who did not make it to the campus. This is enough to expand the enrolment by at least twice the current enrolment in CLC. 
2. Since 2008, Main Campus's passers in the top 1620 exceeded the available slots by an average of 625 NCE passers who did not make it to PSHS. This means that Main Campus can choose to open as much as four times its current enrolment as much as its funding, facilities and availability of faculty members allow.
3. Since 2008, an average of 39 NCE passers who belong to the top 1620 NCE passers did not get a scholarship slot in SMC. This means that SMC can offer at least one more section of 30 scholarship slots.
4. Since 2008, WVC has an average of 31 NCE passers who belong to the top 1620 NCE passers who did not get a scholarship slot.
5. CBZRC still has 103 NCE passers that belong to the top 1620 NCE passers in 2017 who did not make it to the campus, even after increasing the scholarship slots to 120.

From Table 4, we can see that every campus has enough alternate qualifier to fill 90 scholarship slots with the exceptions of MRC, SRC, and ZRC.

## Recommendations

The following are the recommendations based on above observations.

* Expand the enrolment in the following campuses:

  - Main Campus
  - CLC
  - SMC
  - WVC

* Intensify efforts to increase the quality and number of applicants for MRC, SRC, and ZRC.
