---
title: "How Many Scholarship Slots are too Many in the PSHS Campuses?"
author: Joseph S. Tabadero, Jr.\thanks{Special Science Teacher IV, Philippine Science
  High School-CAR Campus}
date: "9/6/2018"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    df_print: paged
subtitle: Counting Top Passers, and Comparing the Relative Performances Against Principal
  and Alternate Qualifiers
classoptions: a4paper, 11pt
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning =F, cache = T)
```


```{r}
source("nce_data_structuring.R")
```

## Executive Summary

* At least three indicators can be set to determine the capability of a campus to offer more scholarship slots: (1) infrastructure; (2) the depth of the pool of qualified applicants by campus of choice; and (3) the capability of instructional services, by campus, to bring every accepted scholar up to PSHS standards. In this paper, we are going to focus on indicators 2 and 3. 
* To investigate indicator 2, we shall look at the top 1590 passers nationwide to determine the number of scholars in each campus who belonged to the top 1590 and the number of principal qualifiers in each campus who are not in the top 1590, nationwide. The number 1590 is chosen because that is the sum of slots available if we assume 240 slots for the Main Campus and 90 slots for each of the 15 regional campuses. 
* An *NCE passer* is defined as any NCE applicant whose score in each subject is above the subject mean score for the applicable NCE year. A *principal qualifier* is an NCE passer who either belongs to the top 240 NCE passers nationwide or to the top 90 NCE passers for his or her campus of first choice. We note that a principal qualifier may not belong to the top 1590 NCE passers, nationwide. 
* Additionally, we shall look at the top 3180 passers nationwide. Ideally, the alternate qualifiers should belong to the top 3180 passers nationwide.
* Based on the count of top 1590 passers who signified interest for each campus, **Main Campus, CLC, and CBZRC** have passers that almost double the available scholarship slots, throughout the years. Meanwhile, SMC and WVC can offer 30-60 more scholarship slots.
* Regarding alternate qualifiers, based on the top 3180 passers nationwide (to diminish burden on instruction):
  - BRC can choose from among 50-150 alternates.
  - CARC can choose from among 20-60 alternates.
  - CBZRC can choose from among 100-300 alternates.
  - CLC can choose from among 100-400 alternates.
  - CVC can choose from among 30-90 alternates.
  - CVISC can choose from among 30-60 alternates.
  - EVC can choose from among 39-100 alternates.
  - IRC can choose from among 30-60 alternates.
  - Main Campus can choose from about 500 alternates.
  - SMC can choose from about 150 alternates.
  - WVC can choose from 150-200 alternates.
* On the other hand, CRC, MRC, SRC, and ZRC may opt not to enrol some more alternates in order to lessen the burden of instruction. These campuses should campaign for more applicants.
* In terms of the relative performances in the Readiness Examination of the principal and alternative qualifiers, there is no appreciable difference, with the principal qualifiers only having a slight edge in math, physics and statistics. This may mean two things:
  - The alternative qualifiers can catch up with the principal qualifiers in a level learning field.
  - The instructional services in each campus is capable in bringing every scholar up to par.

## Philippine Science High School Everywhere

With the establishment of the MIMAROPA campus, the Philippine Science High School is now one region away from establishing a campus in every region in the country---as mandated by Republic Act 8496 (otherwise known as "Philippine Science High School (PSHS) System Act of 1997") as amended by Republich Act 9036. This mandate empowers the PSHS System's Board of Trustees to "develop policies for the expansion of enrollment in campuses under the PSHS System" (Section 7 (j)). However, up until CALABARZON Campus increased the number of scholarship slots, there has yet to be a system by which the BOT can determine the readiness of a campus to expand expand enrollment.

We look at three factors affecting the capability of a campus to expand enrollment: (1) available infrastructure and budget; (2) the depth of the pool of qualified applicants by campus of choice; and (3) the capability of instructional services to bring every accepted scholar up to PSHS standards. The first factor is easily ascertained by inspection. In this paper, we investigate factors 2 and 3 and propose indicators to for each.

## Counting Qualified Applicants

While the PSHS System is mandated to provide equal opportunity for each region to have an access to a PSHS campus, RA 8496 also requires standards for it to follow by "offering scholarships to deserving students" (Title I, Sec. 4). This is done through a rigorous selection process starting with the National Competitive Examination (NCE). 

The NCE is composed of 4 tests---verbal, mathematics, science, and abstract reasoning. An *NCE passer* is defined as any NCE applicant whose score in each subject is above the subject mean score for the applicable NCE year. A *principal qualifier* is an NCE passer who either belongs to the top 240 NCE passers nationwide or to the top 90 NCE passers for his or her campus of first choice. 

A natural choice of indicator for the second factor is the number of passers each campus of first choice has in the top 1590 passing examinees nationwide. Here, we note that a principal qualifier in a campus might not be in the top 1590 passers. We arrive at the number 1590 by assuming 240 scholarship slots for Main Campus and 90 scholarship slots for each of the 15 existing regional campuses. Our assumption is that if the quality of applicants are the same throughout the regions of the country, then the proportion of NCE passers in the top 1590 from each region should be the same. This indicator can then show us: (i) how many NCE passers who belonged to the top 1590 did not qualify as a principal qualifier; and, (ii) how many principal qualifiers do not belong to the top 1590 NCE passers. Counting the number of the top 1590 NCE passers in each campus gives us an idea of whether or not we are giving scholarship slots to the most deserving applicants, while maintaining equal opportunity for applicants from the different regions to avail of a scholarship.

The top 3180 NCE passers, on the other hand, can give us an idea on how deep in the list of alternate qualifiers the PSHS can go and still ensure that the quality of scholars is not compromised. Again, we can count the number of passers for each campus of choice who belong to the top 3180 NCE passers.

## How good is the quality of instruction in PSHS?

A critical factor in expanding enrollment is the availability of qualified teachers for each of the regional campuses. The members of the faculty of PSHS are usually degree holders in specialized fields. A measure of the quality of instruction given by teachers is the Readiness Examination (RE) score. The RE is given to Grades 8 and 10 in order to determine the readiness of scholars to pursue more advanced studies in line with the included subjects: integrated science, mathematics, social science, english, statistics, filipino. 

It can been shown that economic status correlates positively with NCE performance. Instruction should be able to improve the readiness of every scholar---principal qualifier or otherwise. Instruction in PSHS should bridge the gap brought by socio-economic background and level the playing field for each scholar. In such an environment, there should only be a few extremely exceptional scholars---all else should perform almost equally well.

A comparison of the distribution of RE scores of principal and alternate qualifiers is therefore required.

## Principal Qualifiers

```{r}
ncetop1590 <- ncetd %>% 
  filter(passflag2 == "passed" | pass_flag == 1) %>%
  group_by(year) %>%
  top_n(1620, score)
```

```{r}
# ncetd %>% tabyl(year, pass_flag, passflag2)
```
```{r}
ncetop1620diff <- ncetop1590 %>% tabyl(choice1, year) %>%
  gather(year, N, 2:8) %>% 
  mutate(Diff = case_when(choice1 == "MC"~N-240,
                          choice1 == "CBZRC"~N-120,
                          TRUE ~ N-90)) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))
```

```{r}
ncetop1620diff %>%
  unite(Diff, N, Diff, sep = ": ") %>%
  spread(year, Diff) %>%
  rename(Campus = choice1, "2011"="2011-01-01", "2012"="2012-01-01", "2013"="2013-01-01", "2014"="2014-01-01", "2015"="2015-01-01", "2016"="2016-01-01", "2017"="2017-01-01") %>%
  filter(!is.na(Campus)) %>%
  kable("latex", booktabs = TRUE, caption = "Number of students belonging to the top 1590 examinees by signified campus of first choice, $N_1$, and the number of students who did not make it as a principal qualifier in their choice campus (positive) or the number of students who did not make it to the top 1590 nationally but made it as principal qualifiers (negative) by region, $N_2$ ($N_1: N_2$).")
```

```{r}
ncetop1620diff %>% filter(!is.na(choice1)) %>% ggplot(aes(year, Diff)) + geom_line() + geom_hline(aes(yintercept = 0), linetype = "dashed") + facet_wrap(~choice1, scales = "free_y") +
  theme_bw()
```

## Alternates


```{r}
ncetop3240 <- ncetd %>% 
  filter(passflag2 == "passed" | pass_flag == 1) %>%
  group_by(year) %>%
  top_n(3180, score)
```

```{r}
# ncetd %>% tabyl(year, pass_flag, passflag2)
```
```{r}
ncetop3240diff <- ncetop3240 %>% tabyl(choice1, year) %>%
  gather(year, N, 2:8) %>% 
  mutate(Diff = case_when(choice1 == "MC"~N-240,
                          choice1 == "CBZRC"~N-120,
                          TRUE ~ N-90)) %>%
  mutate(year = lubridate::ymd(year, truncated = 2L))
```

```{r}
ncetop3240diff %>%
  unite(Diff, N, Diff, sep = ": ") %>%
  spread(year, Diff) %>%
  rename(Campus = choice1, "2011"="2011-01-01", "2012"="2012-01-01", "2013"="2013-01-01", "2014"="2014-01-01", "2015"="2015-01-01", "2016"="2016-01-01", "2017"="2017-01-01") %>%
  filter(!is.na(Campus)) %>%
  kable("markdown", caption = "Number of students belonging to the top 3180 examinees by signified campus of first choice, $N_1$, and the number of students who did not make it as a principal qualifier in their choice campus (positive) or the number of students who did not make it to the top 3180 nationally but made it as principal qualifiers (negative) by region, $N_2$ ($N_1: N_2$).")
```

```{r}
ncetop3240diff %>% filter(!is.na(choice1)) %>% ggplot(aes(year, Diff)) + geom_line() + geom_hline(aes(yintercept = 0), linetype = "dashed") + facet_wrap(~choice1, scales = "free_y") +
  theme_bw()
```

```{r}
#ncetop3180 %>%
#  group_by(year, choice1, id) %>%
#  summarise(total = sum(score)) %>%
#  ggplot(aes(total)) + geom_histogram(bins = 30) + facet_wrap(~choice1)
```


```{r}
# uncomment in final report
# ncetop3180 %>%
#   group_by(choice1, year, id) %>%
#   summarise(total = sum(score)) %>%
#   mutate(pct = total / 220 * 100) %>%
#   summarise_at(vars(pct), funs(min,q1 = quantile(.,0.25), median, mean, q3 = quantile(.,.75),max, sd)) %>%
#   kable("markdown", latex_options = c("striped", "repeat_header"), digits = 2, caption = "Distribution of Total Score Percentages by Campus and by NCE Year")
```


```{r}
re2015_18 <- readit("./Data/processed/re2015_18.csv")
```
```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(campus, score)) + geom_boxplot() + facet_wrap(~grade+subject) + coord_flip()
```

```{r}
# uncomment in final report
# re2015_18 %>% filter(assessment == "re") %>% group_by(year, campus, grade, subject, qualitype, sname) %>%
#   summarise_at(vars(score), sum) %>%
#   ungroup() %>%
#   group_by(year, campus, grade, subject, qualitype) %>%
#   summarise_at(vars(score), funs(min, Q3 = quantile(., .25, na.rm = T), median, Q4 = quantile(., .75, na.rm = T), max, mean(.,na.rm = T), sd(.,na.rm = T))) %>%
#   kable("markdown", latex_options = c("striped", "repeat_header"), digits = 2)
```


```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Bicol Region Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "BRC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


### CAR Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "CARC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### CBZRC Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "CBZRC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Central Luzon Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "CLC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Central Mindanao Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "CMC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Caraga Region Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "CRC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Cagayan Valley Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "CVC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Central Visayas Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "CVISC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Eastern Visayas Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "EVC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Ilocos Region Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "IRC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Main Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "MC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### MIMAROPA Region Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "MRC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Southern Mindanao Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "SMC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### SRC Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "SRC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Western Visayas Campus


```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "WVC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### CAR Campus

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re", campus == "ZRC") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## PSHS System

```{r, fig.height=12, fig.width=7}
re2015_18 %>% filter(assessment == "re") %>%
  group_by(year, campus, grade, subject, qualitype, sname) %>%
  summarise_at(vars(score), sum) %>%
  group_by(year, campus, grade, subject, qualitype) %>%
  summarise_at(vars(score), mean) %>%
  ggplot(aes(qualitype, score)) + geom_boxplot() + facet_wrap(~grade+subject) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

